---
title: "Education"
author: "Aaron Grenz & Marie Marshall"
output: html_document
---

```{r,echo=FALSE, warning=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
library(ggmap)
library(gridExtra)
library(reshape2)
library(tm)
library(wordcloud)
library(stringr)
library(reshape2)
library(scales)
library(googleVis)
library(datasets)
```

```{r,echo=FALSE, warning=FALSE,message=FALSE}

##Load data tables into the global environment.

tab1 = read.csv("C:/Users/Aaron Grenz/Documents/GitHub/Education/Tab 1.csv")
tab2 = read.csv("C:/Users/Aaron Grenz/Documents/GitHub/Education/Tab 2.csv")
tab3 = read.csv("C:/Users/Aaron Grenz/Documents/GitHub/Education/Tab 3.csv")

```

##Percentage of African Americans taking the Exam in American School System for 2013

Arizona had the highest participation rate for African Americans in 2013.  West Virginia had the lowest participation rate for African Americans in 2013. Disclaimer: some states did not have enough data to produce meaningful results.


```{r,echo=FALSE, warning=FALSE,message=FALSE}

blackCleaned0 <- tab2 %>%
  select(1,15) %>%
  filter(Percent.of.Blacks.taking.exam!="0",Percent.of.Blacks.taking.exam!="*",State!="")

ggplot(blackCleaned0,aes(x=reorder(State,-Percent.of.Blacks.taking.exam),y=Percent.of.Blacks.taking.exam)) + geom_bar(stat="identity") + ggtitle("Participation of African American Students") + theme(legend.position="none") + xlab("States") + ylab("% African American taking the test in 2013") + theme_bw() + theme(axis.text.x = element_text(angle=-90))

```

##Histogram of the Percentage of Blacks taking the Exam using GoogleVis

This graph shows there is a higher frequency of schools in the 0 to 2% range.  Some schools in this range include Kansas, California, and Washington. Interestingly, Georgia, Arizona, and Maryland had the highest participation out of all the states.


```{r,echo=FALSE, warning=FALSE,message=FALSE}

blackCleaned1 <- tab2 %>%
  select(1,14) %>%
  filter(Percent.of.Blacks.Passed!="0",Percent.of.Blacks.Passed!="0.00",Percent.of.Blacks.Passed!="*",State!="")

blackCleaned1$Percent.of.Blacks.Passed <- as.numeric(as.character(blackCleaned1$Percent.of.Blacks.Passed))

blackState <- gvisHistogram(blackCleaned1, options=list(width=600, height=400, legend = "none", title="Histogram of % of Blacks Passed", titleTextStyle="{color:'red', fontName:'Courier', fontSize:16}"))

plot(blackState)
print(blackState,file="blackState.html")

```
<iframe src="blackState.html" width="600" height="600"></iframe>

##US Intensity Charts with GoogleVis

This GoogleVis chart shows the percentage of African Americans passing the exam.  The darker the color of a state, the higher the passing percentage.


```{r,echo=FALSE, warning=FALSE,message=FALSE}

blackExam <- gvisGeoChart(blackCleaned1, "State", "Percent.of.Blacks.Passed", options=list(region="US", displayMode="regions", resolution="provinces", width=600, height=400, colors="['#5C3292']"))

plot(blackExam)
print(blackExam,file="blackExam.html")

```
<iframe src="blackExam.html" width="600" height="600"></iframe>

This GoogleVis chart shows the percentage of Hispanic passing the exam.  The darker the color of a state, the higher the passing percentage.

```{r,echo=FALSE, warning=FALSE,message=FALSE}

hispanicCleaned0 <- tab2 %>%
  select(1,23) %>%
  filter(Percent.of.Hispanic.Passed!="0",Percent.of.Hispanic.Passed!="0.00",Percent.of.Hispanic.Passed!="*",State!="")

hispanicCleaned0$Percent.of.Hispanic.Passed <- as.numeric(as.character(hispanicCleaned0$Percent.of.Hispanic.Passed))

hispanicExam <- gvisGeoChart(hispanicCleaned0, "State", "Percent.of.Hispanic.Passed", options=list(region="US", displayMode="regions", resolution="provinces", width=600, height=400, colors="['#871b47']"))

plot(hispanicExam)
print(hispanicExam,file="hispanicExam.html")

```
<iframe src="hispanicExam.html" width="600" height="600"></iframe>

##Interesting Self-Created Graphs

We completed a short comparison on the pass rates of students in Washington and Oregon.  Washington had a significantly higher passing rate compared to Oregon in 2013 for all students. Interestingly enough, the percentage of females who passed the exam only favored the Washington side by 0.05% in 2013.

```{r,echo=FALSE, warning=FALSE,message=FALSE}

##Washington vs. Oregon

comp=tab2 %>%
  filter(State %in% c("Washington","Oregon"))

comp$Percent.Passed <- as.numeric(as.character(comp$Percent.Passed))
comp$Percent.of.Females.Passed <- as.numeric(as.character(comp$Percent.of.Females.Passed))

ggplot(comp,aes(x=State,y=Percent.Passed, fill=State)) + geom_bar(stat="identity") + ggtitle("") + theme(legend.position="none") + ggtitle("Pass Rates: Oregon vs. Washington") + xlab("States") + ylab("% Students Passed") + theme_bw() + theme(axis.text.x = element_text(angle=-90))

ggplot(comp,aes(x=State,y=Percent.of.Females.Passed, fill=State)) + geom_bar(stat="identity") + ggtitle("") + theme(legend.position="none") + ggtitle("Female Pass Rates: Oregon vs. Washington") + xlab("States") + ylab("% Females Passed") + theme_bw() + theme(axis.text.x = element_text(angle=-90))

```

